---
title: "Final Project Report"
author: Saumyaa Sinha, Andrew Harris, Sanjana Kuchipudi
date: "Due: 05/04/2022"
output: html_notebook
---

Note: The csv files are in the github repo, we had to make minor changes to the raw_data.csv file.


links to the data sets below: 
https://www.kaggle.com/code/paultimothymooney/2020-usa-election-vote-percentages-by-state/data?select=democratic_vs_republican_votes_by_usa_state_2020.csv
Shown above is the data set for the democratic vs republican votes data by the US state in 2020. This data is very useful to see the voting percentages in the 2020 presidential election. The variables include the US state and the number of republicans and democrats for each state. Includes additional columns for the US state and the US state code. The last variable is the percent democrat for each of the 50 US states. 

https://www.kaggle.com/datasets/benhamner/2016-us-election?select=county_facts.csv
The 2016 US election data shows the county demographics data for US counties from the US census. This data set has 54 total variables including the fips code, the area name, state abbreviation, population estimates for several years such as 2010 and 2014. It also included the population percent for several age ranges. 

https://www.kff.org/other/state-indicator/distribution-by-age/?currentTimeframe=0&sortModel=%7B%22colId%22:%22Location%22,%22sort%22:%22asc%22%7D
This data set shows the population distributions by age in the US states. It includes the percentages for  six age ranges for all 50 states. This data is based on the Census Bureau’s American Community Survey (ACS). 


Research question: 
How does the age in percent of population affect the voting patterns in the US?


Explains why the topic is important: 
Voter turnout is low in the United states when compared with other similar countries around the world. We want to investigate this turnout, specifically when looking at different age groups, to see if it provides any insight into why our turnout is so low. When looking at different age groups, we can see if efforts being done to improve the voting of those groups are effective. With this being done, it could give some ideas into our problems and help those trying to solve them. 


```{r}
demographics <- read.csv(file.choose("county_facts.csv"))
voting <- read.csv(file.choose("democratic_vs_republican_votes_by_usa_state_2020.csv"))
```

```{r}
raw_data <- read.csv(file.choose("raw_data.csv"))
```

```{r}
raw_data
```
imported datasets

```{r}
library(dplyr)
library(tidyverse)
library(ggplot2)
install.packages("usmap")
library(usmap)
```
install packages / libraries

```{r}
head(demographics)
```



```{r}
glimpse(voting)
```

```{r}
head(voting)
```
```{r}
head(raw_data)
```
looked at all the data and included descriptons of the dataset above

```{r}
newDem <-demographics %>%
  filter(!grepl(".", state_abbreviation, ignore.case = TRUE))
updated <-select(newDem, c('area_name','PST045214','PST120214','AGE135214','AGE295214','AGE775214', 'fips'))
updated
```

filtered out to only get the colums we wanted

```{r}
finalTable <-
  updated %>%
   left_join(voting, by = c("area_name" ="state"))
finalTable
``` 

Joined voting with updated, saved in new table

```{r}
FinalTable2 <-
  finalTable %>%
  left_join(raw_data, by = c( "area_name" = "Location"))
FinalTable2
```
Another join, last one

```{r}
plot_usmap(data = voting, values = "percent_democrat", color = "black") + 
  scale_fill_continuous(name = "Percent of Democrats", label = scales::comma) + 
  theme(legend.position = "right")
```
Plot of U.S with percent of democrats for each state. Darker blue is less democratic. We use this as a baseline understanding of percentage of democrats in each state before looking at age breakdowns. 

```{r}
  FinalTable2 %>%
    ggplot(aes(x = Adults.19.25, y= percent_democrat, color = usa_state)) +
    geom_point(stat = "identity") +
    labs (x = "") 
```
We thought this one would be more democratic than it was, which suprised us. We were expecting a trend to exist with the younger population to be either more democratic or less, but it was evenly spread for this age group.

```{r}
 FinalTable2 %>%
    ggplot(aes(x = Adults.26.34, y= percent_democrat, color = usa_state)) +
    geom_point(stat = "identity") +
    labs (x = "") 
```


```{r}
FinalTable2 %>%
    ggplot(aes(x = Adults.35.54, y= percent_democrat, color = usa_state)) +
    geom_point(stat = "identity") +
    labs (x = "")
```


```{r}
FinalTable2 %>%
    ggplot(aes(x = Adults.55.64, y= percent_democrat, color = usa_state)) +
    geom_point(stat = "identity") +
    labs (x = "") 
```


```{r}
FinalTable2 %>%
    ggplot(aes(x = X65., y= percent_democrat, color = usa_state)) +
    geom_point(stat = "identity") +
    labs (x = "") 
```
These other graphs also follow the same trend as the first one of the ages having little effect on the political affiliation of the populations.


From these graphs of different ages we saw that there was a fairly even spread out which shows us that for the U.S as a whole, this age range doesnt lean towards high or low amount of democrats. 



